---
title: "excercise2"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r cars}
summary(iris)
```

Plot the Sepal.Length vs. Sepal.Width
```{r}
plot(iris$Sepal.Length, iris$Sepal.Width,pch=19, col=iris$Species)
legend("topright", legend = unique(iris$Species), col = unique(iris$Species), pch = 19, title = "Species")

```

<p>
Calculate the covariance of the 4 quantitative columns:

```{r}
iris_quant = iris[, c("Sepal.Length", "Sepal.Width", "Petal.Length", "Petal.Width")]
cov_matrix = cov(iris_quant)
print(cov_matrix)
```
Calculate the correlation of the 4 columns:
```{r}
cor_matrix = cor(iris_quant)
print(cor_matrix)
```
Now only consider the data for Species *setosa*

```{r}
library(dplyr)
iris_setosa = iris %>%
  filter(Species=="setosa")
head(iris_setosa)
```

Also calculate the covariance and correlation matrix for the setosa-only dataset:

```{r}
# get cov matrix
iris_setosa <- iris_setosa[, -5]
cov_mat_setosa = cov(iris_setosa)
print(cov_mat_setosa)
```
```{r}
cor_mat_setosa = cor(iris_setosa)
print(cor_mat_setosa)
```
*Interestingly, there is a positive covariance and correlation between all the variables in the setosa dataset. Whereas, in the big dataset, there were many negative dependencies!*

Why? Because there are different groups in the data. Within the groups, the correlation and covariance is positive. However, when looking at all the points, then there is a negative covariance and correlation. (Like example of Swiss Bank Notes)
<br>
<br>
2.3
I will use the original data and transform the unit from centimeters to inches and then calculate the mean, covariation and correlation matrix:

```{r}
iris_quant_inches <- iris_quant / 2.54
print('Mean of the col values:')
print(colMeans(iris_quant_inches))
```
```{r}
print('covariation:')
print(cov(iris_quant_inches))
```
```{r}
print('correlation:')
print(cor(iris_quant_inches))
```




